---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# nemwebR

<!-- badges: start -->
<!-- badges: end -->

The goal of `nemwebR` is to enable easy downloading of datasets from the [AEMO NEM Web site](https://www.aemo.com.au/energy-systems/electricity/national-electricity-market-nem/data-nem/market-data-nemweb). `nemwebR` has a series of functions for fetching common datasets such as market prices, generator outputs or bids.

`nemwebR` is designed specifically for those looking to access market data but without access to the MMS Data Model typically available to registered market participants. The likely user is someone looking to run ad hoc analyses on NEM data, primarily on historical datasets (> 2 weeks old).

The naming and design of these functions is intended to mirror the naming and conventions of the [MMS Data Model Report](http://nemweb.com.au/Reports/Current/MMSDataModelReport/Electricity/MMS%20Data%20Model%20Report_files/MMS_2.htm). For clarity, only _public_ datasets (or _private; public next day_) are available.

## Notes

"Why does this package exist if I can just download the csv files from NEMWeb?" Great question. Here's why:

* AEMO csv headers and footers are removed so that the data is ready to use in a clean tabular format
* csv files containing multiple datasets are split out into their consituent datasets
* Empty and deprecated columns are stripped out to keep the final data cleaner
* When dealing with dispatch datasets intervention hierarchies have been implemented correctly
* All dates and times are formatted in POSIX +10 timezone (Australia/Brisbane)
* Data types have all been strictly enforced (i.e. dates as POSIX, numeric where they should be)


Depending on the volume of data and update frequency in a given dataset,archive (historical) data usually starts from ~ 2 weeks prior and stretches back to July 2009. Current (recent) data usually stretches back ~ 1 year. The archive functions are typically faster and more efficient than the current functions, so try to use them where possible.


## Five Minute Settlement

As of 1 October 2021 the NEM is dispatched and settled in 5 minute intervals (previously the NEM was dispatched in 5 minute intervals and settled on 30 minute intervals). The old 30 minute interval _trading_ datasets are no longer valid for new data, but exist for historical purposes.  


## Available Datasets - Archive

### Dispatch Data
* DISPATCH_FCAS_REQ -- Regional FCAS requirements
* DISPATCHLOAD -- Scheduled and semi-scheduled generator outputs
* DISPATCHPRICE -- Regional prices
* DISPATCHREGIONSUM -- Regional demand and non-scheduled generation


### Generator Registration data
* DUDETAILSUMMARY -- Summary of registration details for all registered generators
* DUDETAIL -- Registration details for all registered generators
* GENUNITS -- Summary of individual unit registration details, including fuel type and CO~2~ emission information

### Bid Data
* BIDDAYOFFER -- Bid price band data for energy and FCAS markets submitted by generators
* BIDPEROFFER -- Bid generation volumes for energy and FCAS markets submitted by generators
* BIDDUIDDETAILS -- Enablement levels and FCAS trapezium ffor each generator


### Trading Data (Historical 30-minute data)
_Note that these data are no longer updated as of 1 October 2021 due to 5MS_

* TRADINGINTERCONNECT -- Interconnector flows and limits
* TRADINGLOAD -- Scheduled and semi-scheduled generator outputs
* TRADINGPRICE -- Regional prices
* TRADINGREGIONSUM -- Regional demand and non-scheduled generation



## Available Datasets - Current

* DISPATCHPRICE (incomplete, messy)



## Non-NEMWeb data

* Net System Load Profiles -- Annual NSLP data for 2002 onwards



## Installation

Currently only on GitHub, if the package becomes stable enough it might find its way to CRAN.

Install the current version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("aleemon/nemwebR")
```
## Example

A (very) basic example of fetching pricing data across for a given period:

```{r example, warning=FALSE, message=FALSE}
library(nemwebR)
library(tidyverse) # To grab purrr, stringr and the pipe


## Fetch 30-minute pricing for late 2019, early 2020
input <- seq.Date(as.Date("2019-11-01"), as.Date("2020-02-01"), by = "month") %>% str_remove_all("-") %>% str_sub(1, 6)

example_prices <- map_dfr(input, nemwebR_archive_tradingprice)

head(example_prices)

```


## Development

nemwebR is under active development and features may be added or removed. There are number of obvious areas for improvement, but the core functions are sufficient for many analysis applications.



### Future Datasets

* DISPATCHCONSTRAINT
* DISPATCHINTERCONNECTORRES
* DISPATCHCASE_OCD
* CONSTRAINTRELAXATION_OCD
* GENCONDATA
* GENCONSET
* GENCONSETINVOKE
* GENERICCONSTRAINTRHS
* GENERICEQUATIONDESC
* GENERICEQUATIONRHS
* INTERCONNECTOR
* INTERCONNECTORCONSTRAINT
* LOSSFACTORMODEL
* LOSSMODEL
* MARKET_PRICE_THRESHOLDS
* MARKETFEE
* MARKETFEEDATA
* MARKETNOTICEDATA
* MARKETNOTICETYPE
* MTPASA datasates
* P5MIN datasets
* PDPASA datasets
* PREDISPATCH datasets
* RESIDUE datasets
* ROOFTOP_PV_ACTUAL
* ROOFTOP_PV_FORECAST
* STPASA datasets
* TRANSMISSIONLOSSFACTOR
