---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# nemwebR

<!-- badges: start -->
<!-- badges: end -->

The goal of nemwebR is to enable easy downloading of datasets from the [AEMO NEM Web site](https://www.aemo.com.au/energy-systems/electricity/national-electricity-market-nem/data-nem/market-data-nemweb).

nemwebR has a series of functions for fetching common datasets such as market prices, generator outputs or bids.

The naming and design of these functions is intended to mirror the naming and conventions of the [MMS Data Model Report](http://nemweb.com.au/Reports/Current/MMSDataModelReport/Electricity/MMS%20Data%20Model%20Report_files/MMS_2.htm). For clarity, only _public_ datasets (or _private; public next day_) are available.

There are a few key differences relative to the raw AEMO data:
- AEMO csv headers and footers are removed so that the data is ready to use
- Empty and deprecated columns are stripped out to keep the data cleaner
- When dealing with dispatch datasets intervention heirachies have already been implemented correctly
- All dates and times are formatted in POSIX +10 (Australia/Brisbane)

## Installation

Install the current version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("aleemon/nemwebR")
```
## Example

A basic example to fetch a year of pricing data:

```{r example}
library(nemwebR)
library(purrr)

## fetch pricing data for 2020
input <- seq(20200101, 20201201, by = 100)

prices_2020 <- map_dfr(input, nemwebR_archive_tradingprice)

head(prices_2020)

```


## Development

nemwebR is under active development and features may be added or removed. There are number of obvious areas for improvment, but the core functions are sufficient for most applications.

Future improvements:
- Speed improvements --- the functions are currently built on the base `read.csv()` function, `dplyr::read_csv()` isn't usable because of the AEMO header structure. A read function from the Data Table package will likely help improve speed
- Formatting challenge ---- Current reports are released based on AEMO trading days (4:30 - 4:00 AM) whereas archive datasets are stored based on settlement (i.e. starting and ending at midnight)
